#Regression Models Peer Assessment 
##Summary
This study is supposed to explore the work for Motor Trend, a magazine about the automobile industry. Looking at a data 
set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per
gallon (MPG) (outcome). They are particularly interested in the following two questions
1.Is an automatic or manual transmission better for MPG?
2.Quantify the MPG difference between automatic and manual transmissions.
##Data Processing
```{r}
mtcars dataset
data(mtcars)
str(mtcars)
```
 'data.frame':   	 32 obs. of  11 variables:
   $ mpg : num 	 21 21 22.8 21.4 18.7 18.1 14.3 24.4 22.8 19.2 ...
   $ cyl : num 		 6 6 4 6 8 6 8 4 4 6 ...
   $ disp: num 	 160 160 108 258 360 ...
   $ hp  : num  	110 110 93 110 175 105 245 62 95 123 ...
   $ drat: num		 3.9 3.9 3.85 3.08 3.15 2.76 3.21 3.69 3.92 3.92 ...
   $ wt  : num 		2.62 2.88 2.32 3.21 3.44 ...
    $ qsec: num  	16.5 17 18.6 19.4 17 ...
   $ vs  : num  		0 0 1 1 0 1 0 1 1 1 ...
   $ am  : num  	1 1 1 0 0 0 0 0 0 0 ...
   $ gear: num  	4 4 4 3 3 3 3 4 4 4 ...
   $ carb: num  	4 4 1 1 2 1 4 2 2 4 ...

The variables present in the dataset are:
mpg - Miles/(US) gallon
cyl - Number of cylinders
disp - Displacement (cu.in.)
hp - Gross horsepower
drat - Rear axle ratio
wt - Weight (lb/1000)
qsec - 1/4 mile time
vs - V/S
am - Transmission (0 = automatic, 1 = manual)
gear - Number of forward gears
carb - Number of carburators
Let's convert am feature into a categorical one:
```{r}
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <-c("AT", "MT") ## AT for Automatic T## MT for Manual 
```
Exploratory Data Analysis
First, we need to check if mpg has a distribution close to the gaussian distribution so we can apply regression.
The mpg histogram is presented in Appendix.1 and in spite of the reduced observations in mtcars definitely it resembles 
a normal distribution.
Data Analysis
Next, the means and ranges of mpg for AT versus MT are presented in Appendix.2 boxplot
“Is an automatic or manual transmission better for MPG ?”
To get exact values and confidence interval for fuel consumption by AT vs. MT vehicles, we split the dataset by AT vs. 
MT
and apply the t-test:
```{r}
mpg.at <- mtcars[mtcars$am == "AT",]$mpg
mpg.mt <- mtcars[mtcars$am == "MT",]$mpg
t.test(mpg.at, mpg.mt) 
```
 
  Welch Two Sample t-test
 
 data:  mpg.at and mpg.mt
 t = -3.767, df = 18.33, p-value = 0.001374
 alternative hypothesis: true difference in means is not equal to 0
 95 percent confidence interval:
  -11.28  -3.21
 sample estimates:
 mean of x mean of y 
     17.15     24.39
As the p-value is 0.001374 well bellow 5% or 1%, the alternative hypothesis is true: the difference in means is not 
equal to 0. So, the mean mileage of automatic transmission is 17.15 mpg and the manual transmission is 24.39 mpg.
The 95% confidence interval of the difference in mean gas mileage is between 3.21 and 11.28 mpg.
Regression Analysis
Let's start with the simplest regression model for mpg with only one predictor: am.
```{r}
model1 <- lm(mpg ~ am, data = mtcars)
summary(model1)
```
 
 Call:	
 lm(formula = mpg ~ am, data = mtcars)
# 
 Residuals:
    Min    	 1Q Median   	  3Q    	Max 
 -9.392 	-3.092 -0.297 	 3.244 	 9.508 
 
 Coefficients:
             	Estimate Std. 	Error      t  value  Pr(>|t|)    
 (Intercept)    17.15        	1.12   	15.25  	1.1e-15 ***
  amMT          	  7.24       	1.76   	 4.11 	 0.00029 ***
 ---
 Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
 
 Residual standard error: 4.9 on 30 degrees of freedom
 Multiple R-squared:  0.36,   Adjusted R-squared:  0.338 
 F-statistic: 16.9 on 1 and 30 DF,  p-value: 0.000285
The null hypothesis is rejected by p-value = 0.000285 but the regression model covers only 36% of the variance.
So, the model coefficients are:
intercept = 17.15 represents automatic cars mean mpg
am coefficient = 7.24 represents the adjusted estimate for the expected change in mpg comparing AT versus MT.
Next, we will use a more complex regression model:
```{r}
model.all <- lm(mpg ~ ., data=mtcars)
and go backwards to the fitest model using step() function:
model.best <- step(model.all, trace=0)
summary(model.best)
```
 
 Call:
 lm(formula = mpg ~ wt + qsec + am, data = mtcars)
 
 Residuals:
    Min    	 1Q Median  	   3Q   	 Max 
 -3.481	 -1.556 -0.726 	 1.411  	4.661 
 
 Coefficients:
             	Estimate Std. 	Error    t   value	 Pr(>|t|)    
 (Intercept)       9.618  	  	 6.960    1.38  	0.17792    
 wt          	  -3.917     	 0.711   -5.51    	7e-06 ***
 qsec           	    1.226      	0.289     4.25 	 0.00022 ***
 amMT           	    2.936      	1.411    2.08   	 0.04672 *  
 ---
 Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
 
 Residual standard error: 2.46 on 28 degrees of freedom
 Multiple R-squared:  0.85,   Adjusted R-squared:  0.834 
 F-statistic: 52.7 on 3 and 28 DF,  p-value: 1.21e-11
The “Occam's razor” model explains 85% of mpg variance and contains only 3 predictors:
formula = mpg ~ wt + qsec + am
amMT estimated coefficient equals now to 2.9358 and represents the adjusted estimate for the expected change in mpg 
comparing AT versus MT for this new model containing 2 other predictors besides am.
-amMT estimated coefficient is the answer to the second question.
Best model residuals are depicted in Appendix.3
First graphic, “Residuals vs. Fitted values” is not quite a straight line, proof of some outliers.
Appendix.1 Miles per Gallon Histogram
```{r}
hist(mtcars$mpg, breaks=12, xlab="miles per gallon", main="mpg histogram", col="green")
```
 
Appendix.2 Automatic vs Manual Transmission MPG
```{r}
boxplot(mpg ~ am, data=mtcars, xlab="Transmission Type", ylab="Miles per Gallon",
        main="Automatic versus Manual Transmission MPG")
``` 
 
Appendix.3 Best model residuals
```{r}
par(mfrow = c(2,2))
plot(model.best)
```
 
Final conclusion
From the selected model (fit2), we conclude that the cars with manual transmission have MPG 2.94 higher than the cars 
with automatic transmission.


